\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage{calc}
\usepackage{setspace}
\usepackage{multicol}
\usepackage{fancyheadings}

\usepackage{graphicx}
\usepackage{color}
\usepackage{rotating}
\usepackage{harvard}
\usepackage{aer}
\usepackage{aertt}
\usepackage{verbatim}

\setlength{\voffset}{0in}
\setlength{\topmargin}{0pt}
\setlength{\hoffset}{0pt}
\setlength{\oddsidemargin}{0pt}
\setlength{\headheight}{0pt}
\setlength{\headsep}{.4in}
\setlength{\marginparsep}{0pt}
\setlength{\marginparwidth}{0pt}
\setlength{\marginparpush}{0pt}
\setlength{\footskip}{.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\parskip}{0pc}

\renewcommand{\baselinestretch}{1.5}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}
\newcommand{\prbf}[1]{\textbf{#1}}
\newcommand{\prit}[1]{\textit{#1}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bdm}{\begin{displaymath}}
\newcommand{\edm}{\end{displaymath}}
\newcommand{\script}[1]{\begin{cal}#1\end{cal}}
\newcommand{\citee}[1]{\citename{#1} (\citeyear{#1})}
\newcommand{\h}[1]{\hat{#1}}
\newcommand{\ds}{\displaystyle}

\newcommand{\app}
{
\appendix
}

\newcommand{\appsection}[1]
{
\let\oldthesection\thesection
\renewcommand{\thesection}{Appendix \oldthesection}
\section{#1}\let\thesection\oldthesection
\renewcommand{\theequation}{\thesection\arabic{equation}}
\setcounter{equation}{0}
}

\pagestyle{fancyplain}
\lhead{}
\chead{Empirical Significance of Learning with Firm-Specific Capital}
\rhead{\thepage}
\lfoot{}
\cfoot{}
\rfoot{}

\begin{document}

\begin{titlepage}
\begin{singlespace}
\title{Empirical Significance of Learning in a New Keynesian Model with Firm-Specific Capital\footnote{I am grateful for the advice and guidance of Eric Leeper, Kim Huynh, Brian Peterson, and Todd Walker; for useful conversations with Pedro Falc\~{a}o de Araujo, James Bullard, Troy Davig, Kenneth Kasa, Fabio Milani, Michael Plante, and Bruce Preston; and for comments by the participants of the 2007 Federal Reserve Bank of St. Louis Learning Week Conference, the 2007 Missouri Economics Conference and Indiana University economics department seminars.  All errors are my own.}}
\date{September 30, 2008}
\author{James Murray\\
Dahl School of Business\\
Viterbo University\footnote{\textit{Mailing address}: 900 Viterbo Drive, La Crosse, WI  54601. \textit{E-mail address}: jmmurray@viterbo.edu.  \textit{Phone number}: (608)738-5408. }}

\maketitle

\thispagestyle{empty}

\abstract{\input{capital1_abstract.tex}} \newline 

\noindent \textit{Keywords}: Learning, firm-specific capital, New Keynesian model, maximum likelihood. \\
\noindent \textit{JEL classification}: C13, E22, E31, E50.
\end{singlespace}
\end{titlepage}
\newpage

\input{capital1_body.tex}
\newpage
\input{capital1_bib.tex}
\newpage
\input{capital1_app.tex}
\input{capital1_tabsfigs.tex}

\end{document}



